---
output:
  md_document:
    variant: markdown_github
---

[![Build Status](https://travis-ci.com/jkbye/simpr.svg?branch=master)](https://travis-ci.com/statisfactions/simpr)

# simpr

`simpr` provides a simple and tidyverse-friendly syntax for specifying and generating simulations, especially for power analysis.  The primary workflow is:

  1. Specify **variables** for your simulation data, with `variables()`
  2. Specify **parameters** that you want to systematically vary between different cells of your simulation design (e.g. *n*, effect size, distribution type), with `meta()`
  3. Generate the simulation data with `gen()`
  4. Fit models with your data (e.g. `lm()`), with `fit()`
  5. Tidy the model output for further processing, such as computing power or Type I Error rates, with `calc_tidy`
  
## Installation

```{r eval = FALSE}
remotes::install_github("statisfactions/simpr")

```


## Introductory example: regression with three variables

Let's say we want to see our power to detect an interaction in a linear model.   We can completely run the simulation and calculate *p* values with a just a few lines of code:
```{r}
library(simpr)

set.seed(100)
simpr_tidy = ## Specify the simulation
  variables(x1 = ~ 2 + rnorm(n),
                       x2 = ~ 3 + 2*x1 + rnorm(n, 0, sd = 0.5),
                       y = ~ 5 + b1*x1 + b2*x2 + g1*x1*x2 + 10 * rnorm(n)) %>%
  meta(n = seq(100, 300, by = 20),
       b1 = 1,
       b2 = 1,
       g1 = seq(-1, 1, by = 0.5)) %>% 
  ## Generate the data
  gen(10) %>% 
  ## Fit models
  fit(lm = ~lm(y ~ x1*x2, data = .)) %>% 
  ## Calculate the output
  calc_tidy

```

This gives a tibble with slope estimates and *p* values which we can use to plot the power curves:

```{r message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)

simpr_tidy %>%
  filter(term %in% "x1:x2") %>%
  group_by(n, g1) %>%
  summarize(power = mean(p.value < 0.05)) %>%
  ggplot(aes(n, power)) +
  geom_line() +
  facet_grid(~g1)
```

### Breaking down the example

First, we specify how we want the data to be generated:

```{r}

simpr_spec = variables(x1 = ~ 2 + rnorm(n),
                       x2 = ~ 3 + 2*x1 + rnorm(n, 0, sd = 0.5),
                       y = ~ 5 + b1*x1 + b2*x2 + g1*x1*x2 + 10 * rnorm(n)) %>%
  meta(n = seq(100, 300, by = 20),
       b1 = 1,
       b2 = 1,
       g1 = seq(-1, 1, by = 0.5))
```

The call to `variables()` contains the basics of what we actually want simulated.  Each argument is a named, one-sided formula that can include functions like `rnorm` or whatever else you want, specified similar to `purrr` formula functions.  Note that these arguments include both references to previously defined variables (`x1` and `x2`), and to some other variables not yet defined (`n`, the sample size; `b1`, the slope of `x1`; `b2`, the slope of `x2`; `g1`, the interaction slope).

We can define these variables, which we call *metaparameters* of the simulation, in the `meta()` command.  `meta()` also takes named arguments, and here we define what those metaparameters are.  We can specify them either as constants, or as lists or vectors; `simpr` will generate all possible combinations of these metaparameters and run the simulation for each combination using `gen()`:

```{r}
simpr_gen = simpr_spec %>% 
  gen(10)
```

`gen` has one argument, the number of repetitions for each simulation.  Here we generate 10 repetitions. This produces a `tibble` with one row for each combination of metaparameters and repetition, and a list-column with the generated data.

```{r}
simpr_gen
```

Note the first 5 rows have everything the same with the column `n`, but `g1` varies, and each element of `sim_cell` is a tibble with 100 rows; then on the sixth row, we have the next value of `n`, 120, and a tibble with 120 rows, and so on. Each element of  `sim_cell` contains the generated `x1`, `x2`, and `y`, e.g.

```{r}
simpr_gen$sim_cell[[1]]
```

Next, we can fit a model on this data using the `fit()` function; this uses similar formula syntax to `variables()`:

```{r}
simpr_fit = simpr_gen %>% 
  fit(lm = ~lm(y ~ x1*x2, data = .))
```

Using `purrr` syntax, we refer to each simulated dataset that we want to fit the model to as `.`.

This just adds a list-column onto `simpr_gen` with the model fit for each rep and metaparameter combination:

```{r}
simpr_fit
```

But we can simplify this a lot more for the purposes of power analysis by using `calc_tidy()`, which runs `broom::tidy()` on each of the `lm` objects and brings everything together into one data frame:

```{r}
simpr_tidy = simpr_fit %>% 
  calc_tidy

simpr_tidy
```

This gives a data frame with one row for each term for each combination of metaparameters.  We can easily plot power to detect the interaction term using `dplyr` and `ggplot2`, as noted above, and these commands can be easily chained.
