---
title: "getting-started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(simpr)
set.seed(2001)
```

This vignette provides an overview of each step in the `simpr` workflow:

  1. Specify your data-generating process for each of your variables with `specify()`
  2. Define parameters that you want to systematically vary between different cells of your simulation design (e.g. *n*, effect size), with `define()`
  3. Generate the simulation data with `generate()`
  4. Fit models with your data (e.g. `lm()`), with `fit()`
  5. Tidy the model output for further processing, such as computing power or Type I Error rates, with `tidy_fits`
  
## `specify()` your data-generating process

`specify()` takes arbitrary R expressions that can be used for generating data.  **Each argument should have a name and be prefixed with `~`, the tilde operator.**  Order matters: later arguments can refer to earlier arguments, but not the other way around.

Good:
```{r correct_order}
specify(a = ~ runif(6), 
        b = ~ a + rnorm(6)) %>% 
  generate(1)

```

Error:
```{r incorrect_order, warning = TRUE, purl = FALSE}
specify(b = ~ a + rnorm(6),
        a = ~ runif(6)) %>% 
  generate(1)
```

All arguments must imply the same number of rows.  Arguments that imply 1 row are recycled.

OK:
```{r correct_number}
specify(a = ~ runif(6), 
        b = ~ rnorm(6)) %>% 
  generate(1)
```

OK, with recycling:

```{r recycle_number}
specify(a = ~ runif(1), 
        b = ~ rnorm(6)) %>% 
  generate(1)
```

Error:
```{r incorrect_number, warning = TRUE, purl = FALSE}
specify(a = ~ runif(2), 
        b = ~ rnorm(6)) %>% 
  generate(1)
```

Using `x` as an argument to `specify()` is not recommended, because for technical reasons `x` is always placed as the first argument.  This means that if `x` refers to prior variables it will fail:

```{r x_error, warning = TRUE, purl = FALSE}
specify(y = ~ runif(6),
        x = ~ y + runif(6)) %>% 
  generate(1)

```

The same code works fine when `x` is renamed:

```{r x_error_fixed}
specify(y = ~ runif(6),
        a = ~ y + runif(6)) %>% 
  generate(1)
```

### Expressions that generate multiple columns
`specify()` accepts expressions that generate multiple columns simultaneously in a `matrix`, `data.frame`, or `tibble`.  By default, the column names in the output append a number to the variable name.  Here's an example using `MASS::mvrnorm()`, which produces draws form the multivariate normal distribution:

```{r multicolumn_default}
specify(a = ~ MASS::mvrnorm(6, 
                            mu = rep(0, 3),
                            Sigma = diag(rep(1, 3)))) %>% 
  generate(1)

```

You can change the separator via `sep`:
```{r multicolumn_sep}
specify(a = ~ MASS::mvrnorm(6, 
                            mu = rep(0, 3),
                            Sigma = diag(rep(1, 3))),
        sep = ".") %>% 
  generate(1)

```

Alternatively, you can set a two-sided formula to set names; the left hand side must use `c` or `cbind` and the argument name is ignored:

```{r multicolumn_two_sided}
specify(y = c(a, b, c) ~ MASS::mvrnorm(6, 
                            mu = rep(0, 3),
                            Sigma = diag(rep(1, 3)))) %>% 
  generate(1)

```

If your expression already produces column names, those are used by default and the argument name is ignored:

```{r multicolumn_use_names}
specify(a = ~ MASS::mvrnorm(6, 
                            mu = rep(0, 3),
                            Sigma = diag(rep(1, 3))) %>% 
          magrittr::set_colnames(c("d", "e", "f"))) %>% 
  generate(1)
```

This is useful for dealing with functions from other packages that already provide informative names (e.g., `lavaan::simulateData()`). You can turn this behavior off with `use_names = FALSE`.

Whatever method you use, you can still refer to these generated variables in later expressions:

```{r multicolumn_refer}
specify(a = ~ MASS::mvrnorm(6, 
                            mu = rep(0, 3),
                            Sigma = diag(rep(1, 3))),
        b = ~ a_1 - a_2) %>% 
  generate(1)
```

## `define()` parameters that you want to systematically vary

## `generate()` the simulation data

## `fit()` models with your data

## `tidy_fits()` for further processing



